---
title: "Household SDSE and S. pyogenes co-occurrence"
author: "Ouli Xie"
date: "07/06/2023"
output: pdf_document
---

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggsci)
library(svglite)
```

## Evaluate frequency of co-occurrence of SDSE and S. pyogenes in households

Takes into account all positive swabs, including those which were not sequenced.
To generate model of independent transmission (both across and within species), 
clustering of isolates from the same transmission cluster (inferred using whole 
genome sequencing) within households must be accounted for. Otherwise, 
permutation of these swabs across all swabs taken at a community visit will 
cause separation of the linkage
between these isolates and predict co-occurrences more frequently than would be
expected.

Transmission cluster data was available for 89% of SDSE isolates and 96% of
S. pyogenes isolates. Those without sequencing data were assumed to be separate
transmission clusters.

First read in dataframe with all positive swabs then collapse clustered swabs
(same transmission cluster, same household, same community visit).

```{r, echo=TRUE, results=FALSE, message=FALSE}
all_pos <- read_csv("all_pos_df.csv")

all_pos <- all_pos[is.na(all_pos$gene_and_SNP_cluster) | 
                     !duplicated(all_pos %>% 
                                   dplyr::select(House, gene_and_SNP_cluster, 
                                                 Interval)), ]

# Collapses 90 positives
```

Assign if each person was positive for SDSE and/or pyo at each visit using the
above "collapsed" list of positives. This is a binary yes/no designation and
no longer requires transmission cluster data.

```{r, echo=TRUE, results=FALSE, message=FALSE}
# Read in all swabs
all_swabs <- read_csv("co_occur_df.csv")

all_pos <- all_pos %>% 
  mutate(person_date = paste(QryPerson_2016.Person_id, SpecDate, sep = "_"))
all_swabs <- all_swabs %>% 
  mutate(person_date = paste(QryPerson_2016.person_ID, Date, sep = "_"))

all_swabs$Date <- as.Date(all_swabs$Date, "%d-%b-%y")
all_pos$SpecDate <- as.Date(all_pos$SpecDate, "%d-%b-%y")

# Check if positive for any SDSE
sdse <- tibble(sdse = rep(0, nrow(all_swabs)), 
               sdse_throat = rep(0, nrow(all_swabs)), 
               sdse_skin = rep(0, nrow(all_swabs)))
for (i in 1:nrow(all_swabs)) {
  if (sum(c("C", "G") %in% 
          all_pos$Strep_Group[which(all_pos$person_date == all_swabs$person_date[i] & 
                                    all_pos$Site == "Throat")]) > 0) {
    sdse[i, 1] <- 1
    sdse[i, 2] <- 1
  }
  if (sum(c("C", "G") %in% 
          all_pos$Strep_Group[which(all_pos$person_date == all_swabs$person_date[i] &
                                    all_pos$Site != "Throat")]) > 0) {
    sdse[i, 1] <- 1
    sdse[i, 3] <- 1
  }
}
# Check if positive for any pyo
pyo <- tibble(pyo = rep(0, nrow(all_swabs)), 
              pyo_throat = rep(0, nrow(all_swabs)), 
              pyo_skin = rep(0, nrow(all_swabs)))
for (i in 1:nrow(all_swabs)) {
  if ("A" %in% all_pos$Strep_Group[which(all_pos$person_date == all_swabs$person_date[i] &
                                         all_pos$Site == "Throat")]) {
    pyo[i, 1] <- 1
    pyo[i, 2] <- 1
  }
  if ("A" %in% all_pos$Strep_Group[which(all_pos$person_date == all_swabs$person_date[i] &
                                         all_pos$Site != "Throat")]) {
    pyo[i, 1] <- 1
    pyo[i, 3] <- 1
  }
}

# Add to all_swabs dataframe
all_swabs <- bind_cols(all_swabs, sdse, pyo)
```

```{r, include=FALSE}
# Manually fix some missing values due to date discrepancies (all skin)
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 213 & all_swabs$Date == as.Date("07/12/2004", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 213 & all_swabs$Date == as.Date("07/12/2004", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 1 & all_swabs$Date == as.Date("05/10/2004", "%d/%m/%Y")), 
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 1 & all_swabs$Date == as.Date("05/10/2004", "%d/%m/%Y")), 
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 320 & all_swabs$Date == as.Date("21/04/2004", "%d/%m/%Y")), 
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 320 & all_swabs$Date == as.Date("21/04/2004", "%d/%m/%Y")), 
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 389 & all_swabs$Date == as.Date("06/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 389 & all_swabs$Date == as.Date("06/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 570 & all_swabs$Date == as.Date("18/01/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 570 & all_swabs$Date == as.Date("18/01/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 654 & all_swabs$Date == as.Date("23/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 654 & all_swabs$Date == as.Date("23/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 665 & all_swabs$Date == as.Date("07/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 665 & all_swabs$Date == as.Date("07/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 893 & all_swabs$Date == as.Date("14/02/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 893 & all_swabs$Date == as.Date("14/02/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo_skin")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 1109 & all_swabs$Date == as.Date("21/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo")] <- 1
all_swabs[which(all_swabs$QryPerson_2016.person_ID == 1109 & all_swabs$Date == as.Date("21/06/2005", "%d/%m/%Y")),
          which(colnames(all_swabs) == "pyo_skin")] <- 1
```

Permute SDSE and pyo positive designations across all swabs taken at each
community visit.

The manuscript performs 10,000 permutations but for speed, this script performs
1,000.

Can change count(pyo/sdse) to count(pyo_throat/sdse_throat) for sensitivity
analysis when restricting overlaps to throat level only.

```{r}
permute_interval = tibble(perm_no = 1:100)
permute_interval[, as.character(1:35)] <- 
  rep(0, 100)

# Get total number of co-occurrences at each sampling interval
# Iterate through intervals
for (n in 1:100) {
  for (i in unique(1:35)) {
    swab_interval <- all_swabs %>% filter(Interval == i)
    # Permute the positives across the swabs within the interval
    pyo <- sample(swab_interval$pyo, replace = FALSE)
    sdse <- sample(swab_interval$sdse, replace = FALSE)
    # Count co-occurrences - a co-occurrence is counted if SDSE and pyo are
    # found in a household at the same community visit regardless of if it was
    # in the same individual or across different individuals in a household
    swab_interval$pyo <- pyo
    swab_interval$sdse <- sdse
    pyo_pos <- swab_interval %>% group_by(House) %>% 
      count(pyo) %>% filter(pyo == 1) %>% pull(House)
    sdse_pos <- swab_interval %>% group_by(House) %>% 
      count(sdse) %>% filter(sdse == 1) %>% pull(House)
    permute_interval[n, which(
      colnames(permute_interval) == as.character(i))] <- 
      length(intersect(pyo_pos, sdse_pos))
    permute_same_swab[n, 2] <- permute_same_swab[n, 2] + nrow(swab_interval %>% filter(sdse == 1, pyo == 1))
  }
}
```

Number of observed overlapping household-visits

```{r}
real_interval = tibble()
real_interval[, as.character(unique(sample_intervals$Interval))] <- 0
for (i in unique(sample_intervals$Interval)) {
  swab_interval <- all_swabs %>% filter(Interval == i)
  pyo_pos <- swab_interval %>% group_by(House) %>% 
    count(pyo) %>% filter(pyo == 1) %>% pull(House)
  sdse_pos <- swab_interval %>% group_by(House) %>% 
    count(sdse) %>% filter(sdse == 1) %>% pull(House)
  real_interval[1, which(colnames(real_interval) == as.character(i))] <- length(intersect(pyo_pos, sdse_pos))
}

# Total overlaps in real data
sum(t(real_interval))
```

Plot permuted co-occurrences

```{r}
permute_summary <- permute_interval %>% rowwise() %>% 
  mutate(total = sum(c_across(`1`:`35`))) %>% dplyr::select(total)

permute_plot_all <- permute_summary %>% ggplot(aes(x = total)) + 
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = sum(t(real_interval)), linetype = "dashed", 
             colour = "firebrick3", size = 1) + 
  xlab(label = "Number of co-occurrences") +
  ylab(label = "Number of permutations") + theme_bw(base_size = 20)
```

```{r, echo=FALSE}
permute_plot_all
```

One-sided p value <= observed number of co-occurrences (hypothesis that there is
inter-species transmission interference)

```{r}
nrow(permute_summary %>% filter(total <= 100))/nrow(permute_summary)
```

