---
title: "SDSE_transmission_networks_comm_1"
author: "Ouli Xie"
date: "06/06/2023"
output: pdf_document
---

```{r, include=FALSE}
library(igraph)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(plotfunctions)
library(ggraph)
library(tidygraph)
library(scatterpie)
library(lubridate)
```

## Generate community 1 transmission networks

Read in transmission cluster data for community 1

```{r, echo=TRUE, results=FALSE, message=FALSE}
sdse_df <- read_csv("sdse_clusters_comm_1.csv")
sdse.number.isolates.t.cluster.refined <- nrow(sdse_df)

pyo_df <- read_csv("pyo_clusters_comm_1.csv")
# Remove singletons
pyo_singletons <- pyo_df %>% group_by(t.cluster.refined) %>% count() %>% 
  dplyr::filter(n == 1)
pyo_df <- pyo_df %>% 
  dplyr::filter(!t.cluster.refined %in% pyo_singletons$t.cluster.refined)
pyo.number.isolates.t.cluster.refined <- nrow(pyo_df)

# Create vector of household IDs for transmission matrix - use SDSE order
hh.id <- unique(sdse_df$hh) # order of households in transmission matrix
number.hh <- length(hh.id) # number of households
number.hh.t.cluster.refined <- number.hh
```

Read in sampling interval data

```{r, echo=TRUE, results=FALSE, message=FALSE}
sample_intervals <- read_csv("sampling_intervals.csv")

sample_intervals$Start <- as.Date(sample_intervals$Start, "%d/%m/%Y")
sample_intervals$End <- as.Date(sample_intervals$End, "%d/%m/%Y")
sample_intervals_long <- sample_intervals[0, 1:3] %>% rename(Date = Start)
for (i in 1:nrow(sample_intervals)) {
  for (j in 0:(sample_intervals$Within_interval[i] - 1)) {
    sample_intervals_long <- sample_intervals_long %>% 
      add_row(Community = sample_intervals$Community[i],
              Interval = sample_intervals$Interval[i],
              Date = sample_intervals$Start[i] + j)
  }
}

# Include only Community 1
sample_intervals_long <- sample_intervals_long %>% dplyr::filter(Community == 1)
no_intervals <- length(unique(sample_intervals_long$Interval))
```

Set transmission window

```{r}
# Set transmission window for SDSE (days)
sdse.t.window <- 44
sdse.m.window <- 12

# Set transmission window for pyo (days)
pyo.t.window <- 44
pyo.m.window <- 12
```

Create adjacency matrices for each sampling interval

```{r, echo=TRUE, results=FALSE, message=FALSE}
sdse.t.cluster.refined <- sdse_df
sdse.transmission.determined <- 
  matrix(0, nrow = sdse.number.isolates.t.cluster.refined, ncol = 1)
sdse.mat.t.cluster.refined <- 
  matrix(0, nrow = number.hh, ncol = number.hh)
sdse.mat.t.cluster.refined.carriage <- 
  matrix(0, nrow = number.hh, ncol = number.hh)
rownames(sdse.mat.t.cluster.refined) <- hh.id
colnames(sdse.mat.t.cluster.refined) <- hh.id
rownames(sdse.mat.t.cluster.refined.carriage) <- hh.id
colnames(sdse.mat.t.cluster.refined.carriage) <- hh.id

pyo.t.cluster.refined <- pyo_df
pyo.transmission.determined <- 
  matrix(0, nrow = pyo.number.isolates.t.cluster.refined, ncol = 1)
pyo.mat.t.cluster.refined <- matrix(0, nrow = number.hh, ncol = number.hh)
pyo.mat.t.cluster.refined.carriage <- 
  matrix(0, nrow = number.hh, ncol = number.hh)
rownames(pyo.mat.t.cluster.refined) <- hh.id
colnames(pyo.mat.t.cluster.refined) <- hh.id
rownames(pyo.mat.t.cluster.refined.carriage) <- hh.id
colnames(pyo.mat.t.cluster.refined.carriage) <- hh.id

# Create list for sdse and pyo matrices for each time point
sdse_time_determined <- replicate(no_intervals, sdse.transmission.determined, 
                                  simplify = FALSE)
pyo_time_determined <- replicate(no_intervals, pyo.transmission.determined, 
                                 simplify = FALSE)
sdse_time_matrix <- replicate(no_intervals, sdse.mat.t.cluster.refined,
                              simplify = FALSE)
pyo_time_matrix <- replicate(no_intervals, pyo.mat.t.cluster.refined,
                             simplify = FALSE)
sdse_time_matrix_carriage <- 
  replicate(no_intervals, sdse.mat.t.cluster.refined.carriage,
                                       simplify = FALSE)
pyo_time_matrix_carriage <- 
  replicate(no_intervals, pyo.mat.t.cluster.refined.carriage,
                                      simplify = FALSE)
sdse_time_matrix_skin <- 
  replicate(no_intervals, sdse.mat.t.cluster.refined.carriage,
                                   simplify = FALSE)
pyo_time_matrix_skin <- 
  replicate(no_intervals, pyo.mat.t.cluster.refined.carriage,
                                  simplify = FALSE)
```

Generate SDSE transmission network looking for isolates from the same 
transmission cluster within the transmission window (i.e., between community 
visits). All possible links are drawn

```{r, echo=TRUE, results=FALSE, message=FALSE}
# Read in metadata table to ensure not drawing links between same person
sdse_metadata <- read_csv("sdse_metadata.csv")

for (i in 1:sdse.number.isolates.t.cluster.refined) {
  possiblesource.i <- c()
  # For each isolate j (source)
  for (j in 1:sdse.number.isolates.t.cluster.refined) {
    # That is not isolate i
    if (j!=i) {
      # If isolate i and j belong to the same transmission cluster 
      if (sdse.t.cluster.refined$t.cluster.refined[i]==
          sdse.t.cluster.refined$t.cluster.refined[j]) {
        # Determine date each isolate was collected
        sink.date <- sdse.t.cluster.refined$date[i]
        source.date <- sdse.t.cluster.refined$date[j]
        sink.date <- as.numeric(as.Date(sink.date,format="%d/%m/%Y"))
        source.date <- as.numeric(as.Date(source.date,format="%d/%m/%Y"))
        # Check to ensure not same person
        sink.id <- sdse_metadata$person_id[which(
          sdse_metadata$Strain == sdse.t.cluster.refined$isolate.id[i])]
        source.id <- sdse_metadata$person_id[which(
          sdse_metadata$Strain == sdse.t.cluster.refined$isolate.id[j])]
        if (sink.id != source.id) {
          # if isolate j has a collection date within t.window days of isolate i
          if (sink.date>=source.date+sdse.m.window && 
              sink.date<=source.date+sdse.t.window) { 
            # Add isolate j to list of possible sources of isolate i.  
            possiblesource.i <- append(possiblesource.i,j)
          }
        }
      }
    }
  }
  if (length(possiblesource.i)>0) {
    # Iterate through each possible link and find source and sink intervals
    sink.date <- as.Date(sdse.t.cluster.refined$date[i], "%d/%m/%Y")
    sink.interval <- sample_intervals_long %>% 
      dplyr::filter(Date == sink.date) %>%
      pull(Interval)
    if (sdse_time_determined[[sink.interval]][i,1]==0) {
      for (s in possiblesource.i) {
        source.date <- as.Date(sdse.t.cluster.refined$date[s], "%d/%m/%Y")
        source.interval <- sample_intervals_long %>% 
          dplyr::filter(Date == source.date) %>%
          pull(Interval)
        # Error check
        if (source.date > sink.date) {
          print(i)
          print(s)
          print(source.date)
          print(sink.date)
          stop("Error source date > sink date")
        }
        # Assign links to all possible sources
        # This is the household of sink isolate i
        hh.isolate.i <- sdse.t.cluster.refined$hh[i]
        # This is the position in the matrix of this household
        hh.isolate.i <- which(grepl(hh.isolate.i, hh.id))
        # This is the household of the source isolate
        hh.isolate.j <- sdse.t.cluster.refined$hh[s]
        # This is the position in the matrix of this household
        hh.isolate.j <- which(grepl(hh.isolate.j, hh.id))
        # Add transmission event between these households based on sink interval:
        sdse_time_matrix[[sink.interval]][hh.isolate.j, hh.isolate.i] <-
          sdse_time_matrix[[sink.interval]][hh.isolate.j, hh.isolate.i] + 1
        # If source isolate j is a carriage isolate, add to the carriage matrix
        if (sdse.t.cluster.refined$type[s]=="Aymptomatic_carriage"){
          sdse_time_matrix_carriage[[sink.interval]][hh.isolate.j, hh.isolate.i] <-
            sdse_time_matrix_carriage[[sink.interval]][hh.isolate.j, hh.isolate.i] + 1
        }
        # If source isolate j is a skin isolate, add to the skin matrix
        if (sdse.t.cluster.refined$type[s]=="Impetigo"){
          sdse_time_matrix_skin[[sink.interval]][hh.isolate.j, hh.isolate.i] <-
            sdse_time_matrix_skin[[sink.interval]][hh.isolate.j, hh.isolate.i] + 1
        }
      }
      # Tag isolate as transmission source being determined
      sdse_time_determined[[sink.interval]][i,1] <- 1
    }
  }
}

```

Generate S. pyogenes transmission adjacency matrix looking for isolates from the 
same transmission cluster within the transmission window (i.e., between 
community visits). All possible edges are drawn.

```{r, echo=TRUE, results=FALSE, message=FALSE}
pyo_metadata <- read_csv("pyo_metadata.csv")

for (i in 1:pyo.number.isolates.t.cluster.refined) {
  possiblesource.i <- c()
  # For each isolate j (source)
  for (j in 1:pyo.number.isolates.t.cluster.refined) {
    # That is not isolate i
    if (j!=i) {
      # If isolate i and j belong to the same transmission cluster 
      if (pyo.t.cluster.refined$t.cluster.refined[i] ==
          pyo.t.cluster.refined$t.cluster.refined[j]){
        # Determine date each isolate was collected
        sink.date <- pyo.t.cluster.refined$date[i]
        source.date <- pyo.t.cluster.refined$date[j]
        sink.date <- as.numeric(as.Date(sink.date,format="%d/%m/%Y"))
        source.date <- as.numeric(as.Date(source.date,format="%d/%m/%Y"))
        # Check to ensure not same person
        sink.id <- pyo_metadata$Individual_ID[which(
          pyo_metadata$Isolate_Name == pyo.t.cluster.refined$isolate.id[i])]
        source.id <- pyo_metadata$Individual_ID[which(
          pyo_metadata$Isolate_Name == pyo.t.cluster.refined$isolate.id[j])]
        if (sink.id != source.id) {
          if (sink.date>=source.date+pyo.m.window && 
              sink.date<=source.date+pyo.t.window){ 
            # Add source j to list of possible sources of isolate i.  
            possiblesource.i <- append(possiblesource.i,j)
          }
        }
      }
    }
  }
  if (length(possiblesource.i)>0) {
    # Iterate through each possible link and find source and sink intervals
    sink.date <- as.Date(pyo.t.cluster.refined$date[i], "%d/%m/%Y")
    sink.interval <- sample_intervals_long %>% 
      dplyr::filter(Date == sink.date) %>% pull(Interval)
    if (pyo_time_determined[[sink.interval]][i,1]==0) {
      for (s in possiblesource.i) {
        source.date <- as.Date(pyo.t.cluster.refined$date[s], "%d/%m/%Y")
        source.interval <- sample_intervals_long %>% 
          dplyr::filter(Date == source.date) %>% pull(Interval)
        # Error check
        if (source.date > sink.date) {
          print(i)
          print(s)
          print(source.date)
          print(sink.date)
          stop("Error source date > sink date")
        }
        # Assign links to all possible source
        # This is the household of sink isolate i
        hh.isolate.i <- pyo.t.cluster.refined$hh[i]
        # This is the position in the matrix of this household
        hh.isolate.i <- which(grepl(hh.isolate.i, hh.id))
        # This is the household of the source isolate
        hh.isolate.j <- pyo.t.cluster.refined$hh[s]
        # This is the position in the matrix of this household
        hh.isolate.j <- which(grepl(hh.isolate.j, hh.id))
        # Add transmission event between these households based on sink interval:
        pyo_time_matrix[[sink.interval]][hh.isolate.j, hh.isolate.i] <-
          pyo_time_matrix[[sink.interval]][hh.isolate.j, hh.isolate.i] + 1
        # If the source of isolate is a carriage isolate, add to carriage matrix
        if (pyo.t.cluster.refined$type[s]=="Aymptomatic_carriage"){
          pyo_time_matrix_carriage[[sink.interval]][hh.isolate.j, hh.isolate.i] <-
            pyo_time_matrix_carriage[[sink.interval]][hh.isolate.j, hh.isolate.i] + 1
        } else if (pyo.t.cluster.refined$type[s]=="Impetigo") {
          pyo_time_matrix_skin[[sink.interval]][hh.isolate.j, hh.isolate.i] <-
            pyo_time_matrix_skin[[sink.interval]][hh.isolate.j, hh.isolate.i] + 1
        }
      }
      # Tag isolate as transmission source being determined
      pyo_time_determined[[sink.interval]][i,1] <- 1
    }
  }
}
```

Overlap SDSE and pyo transmission networks

```{r, echo=TRUE, results=FALSE, message=FALSE}
# Read in household size
hh.size <- read_csv("HHsize_community_combined.csv")
# Remove unknown household
hh.size <- hh.size %>% dplyr::filter(hh != 999)
hh.size <- hh.size %>% mutate(hh = paste("HH", hh, sep = "_"))
# Limit to community 1
hh.size <- hh.size %>% dplyr::filter(community == 1)
number.hh <- length(hh.size$hh)

# Read in proportion of positive SDSE and pyo positives in each household
hh_prop <- read_csv("HH_prop_comm_1.csv")

# First take each time point and convert to data frame
union_all_df <- tibble(from = character(), to = character(), 
                       weight_1 = numeric(), weight_2 = numeric(), 
                       color = character(), weight = numeric(), 
                       interval = numeric())
union_throat_df <- tibble(from = character(), to = character(), 
                          weight_1 = numeric(),weight_2 = numeric(), 
                          color = character(), weight = numeric(), 
                          interval = numeric())
for (interval in 1:no_intervals) {
  # Do all isolates first
  # Separate by whether adjacency matrices are occupied by both, one, or none
  if (sum(sdse_time_matrix[[interval]]) > 0 & 
      sum(pyo_time_matrix[[interval]]) > 0) {
    # Undirected weighted graphs
    sdse_net <- graph_from_adjacency_matrix(sdse_time_matrix[[interval]], 
                                            mode="plus", weighted=TRUE, 
                                            diag=TRUE)
    pyo_net <- graph_from_adjacency_matrix(pyo_time_matrix[[interval]], 
                                           mode="plus", weighted=TRUE, 
                                           diag=TRUE)
    # Get union
    union_net <- sdse_net %u% pyo_net
    # Give separate colours if SDSE, pyo or overlapping edges
    for (i in 1:length(E(union_net))) {
      if (!is.na(E(union_net)$weight_1[i]) & !is.na(E(union_net)$weight_2[i])) {
        E(union_net)$color[i] <- "#20854E"
        # Take min weight if overlapping (although use unweighted edges at end)
        E(union_net)$weight[i] <- min(c(E(union_net)$weight_1[i], 
                                        E(union_net)$weight_2[i]))
      } else if (!is.na(E(union_net)$weight_1[i]) & 
                 is.na(E(union_net)$weight_2[i])) {
        E(union_net)$color[i] <- "#0072B5"
        E(union_net)$weight[i] <- min(c(E(union_net)$weight_1[i], 
                                        E(union_net)$weight_2[i]), na.rm = TRUE)
      } else if (is.na(E(union_net)$weight_1[i]) & 
                 !is.na(E(union_net)$weight_2[i])) {
        E(union_net)$color[i] <- "#BC3C29"
        E(union_net)$weight[i] <- min(c(E(union_net)$weight_1[i], 
                                        E(union_net)$weight_2[i]), na.rm = TRUE)
      }
    }
  } else if (sum(sdse_time_matrix[[interval]]) == 0 & 
             sum(pyo_time_matrix[[interval]]) > 0) {
    # Only pyo occupied
    union_net <- graph_from_adjacency_matrix(pyo_time_matrix[[interval]], 
                                             mode="plus", weighted=TRUE, 
                                             diag=TRUE)
    E(union_net)$color <- "#BC3C29"
    # Only SDSE occupied
  } else if (sum(sdse_time_matrix[[interval]]) > 0 & 
             sum(pyo_time_matrix[[interval]]) == 0) {
    union_net <- graph_from_adjacency_matrix(sdse_time_matrix[[interval]], 
                                             mode="plus", weighted=TRUE, 
                                             diag=TRUE)
    E(union_net)$color <- "#0072B5"
  } else {
    # Blank matrix as both empty
    union_net <- graph_from_adjacency_matrix(matrix(0, nrow = number.hh, 
                                                    ncol = number.hh),
                                             mode="plus", weighted=TRUE, 
                                             diag=TRUE)
  }
  # Append to df
  union_net_df <- igraph::as_data_frame(union_net)
  if (nrow(union_net_df) > 0) {
    union_net_df$interval <- interval
  } else if (nrow(union_net_df) == 0) {
    class(union_net_df$from) <- "character"
    class(union_net_df$to) <- "character"
  }
  union_all_df <- union_all_df %>% bind_rows(union_net_df)
  
  # Now do same for throat
  if (sum(sdse_time_matrix_carriage[[interval]]) > 0 & 
      sum(pyo_time_matrix_carriage[[interval]]) > 0) {
    # Undirected weighted graphs
    sdse_throat_net <- graph_from_adjacency_matrix(
      sdse_time_matrix_carriage[[interval]], mode="plus", weighted=TRUE, 
      diag=TRUE)
    pyo_throat_net <- graph_from_adjacency_matrix(
      pyo_time_matrix_carriage[[interval]], mode="plus", weighted=TRUE, 
      diag=TRUE)
    # Get union
    union_throat_net <- sdse_throat_net %u% pyo_throat_net
    # Give separate colours if SDSE, pyo or overlapping edges
    for (i in 1:length(E(union_throat_net))) {
      if (!is.na(E(union_throat_net)$weight_1[i]) & 
          !is.na(E(union_throat_net)$weight_2[i])) {
        E(union_throat_net)$color[i] <- "#20854E"
        # Take min weight although ultimately use unweighted edges
        E(union_throat_net)$weight[i] <- min(c(E(union_throat_net)$weight_1[i], 
                                               E(union_throat_net)$weight_2[i]))
      } else if (!is.na(E(union_throat_net)$weight_1[i]) & 
                 is.na(E(union_throat_net)$weight_2[i])) {
        E(union_throat_net)$color[i] <- "#0072B5"
        E(union_throat_net)$weight[i] <- min(c(E(union_throat_net)$weight_1[i], 
                                               E(union_throat_net)$weight_2[i]), 
                                             na.rm = TRUE)
      } else if (is.na(E(union_throat_net)$weight_1[i]) & 
                 !is.na(E(union_throat_net)$weight_2[i])) {
        E(union_throat_net)$color[i] <- "#BC3C29"
        E(union_throat_net)$weight[i] <- min(c(E(union_throat_net)$weight_1[i], 
                                               E(union_throat_net)$weight_2[i]), 
                                             na.rm = TRUE)
      }
    }
  } else if (sum(sdse_time_matrix_carriage[[interval]]) == 0 & 
             sum(pyo_time_matrix_carriage[[interval]]) > 0) {
    # Only pyo occupied
    union_throat_net <- graph_from_adjacency_matrix(
      pyo_time_matrix_carriage[[interval]], mode="plus", weighted=TRUE, 
      diag=TRUE)
    E(union_throat_net)$color <- "#BC3C29"
    # Only SDSE occupied
  } else if (sum(sdse_time_matrix_carriage[[interval]]) > 0 & 
             sum(pyo_time_matrix_carriage[[interval]]) == 0) {
    union_throat_net <- graph_from_adjacency_matrix(
      sdse_time_matrix_carriage[[interval]], mode="plus", weighted=TRUE, 
      diag=TRUE)
    E(union_throat_net)$color <- "#0072B5"
  } else {
    # Blank matrix as both empty
    union_throat_net <- graph_from_adjacency_matrix(matrix(0, nrow = number.hh, 
                                                           ncol = number.hh),
                                                    mode="plus", weighted=TRUE, 
                                                    diag=TRUE)
  }
  # Append to df
  union_throat_net_df <- igraph::as_data_frame(union_throat_net)
  if (nrow(union_throat_net_df) > 0) {
    union_throat_net_df$interval <- interval
  } else if (nrow(union_throat_net_df) == 0) {
    class(union_throat_net_df$from) <- "character"
    class(union_throat_net_df$to) <- "character"
  }
  union_throat_df <- union_throat_df %>% bind_rows(union_throat_net_df)
}
```

Generate overlapping transmission networks with unweighted edges

```{r}
timed_graph <- as_tbl_graph(union_all_df)
# Use scatterpie to change nodes to pie charts
# Set vertex size
hh.size.temp <- hh.size[hh.size$hh %in% V(timed_graph)$name, ]
hh.size.temp <- hh.size.temp[match(V(timed_graph)$name, hh.size.temp$hh), ]
V(timed_graph)$size <- (hh.size.temp$size)^0.5/100
# Set proportions
hh_prop_network <- hh_prop[match(V(timed_graph)$name, hh_prop$House), ]
V(timed_graph)$pyo <- hh_prop_network$pyo
V(timed_graph)$sdse <- hh_prop_network$SDSE

# Create alternative facet labels
E(timed_graph)$interval2 <- dplyr::recode(E(timed_graph)$interval, 
                                   "7" = "Feb 2004", "8" = "Mar 2004", 
                                   "9" = "Apr 2004", "10" = "May 2004", 
                                   "11" = "Jun 2004", "14" = "Sep 2004", 
                                   "15" = "Oct 2004", "16" = "Nov 2004",
                                   "17" = "Dec 2004", "18" = "Jan 2005", 
                                   "19" = "Feb 2005","20" = "Mar 2005", 
                                   "21" = "Apr 2005", "22" = "May 2005",
                                   "23" = "Jun 2005")

# Precompute layout with FR algorithm
set.seed(60)
l <- layout_with_fr(timed_graph)
V(timed_graph)$x <- l[, 1]
V(timed_graph)$y <- l[, 2]

# Plot unweighted edges
transmission_net <- ggraph(timed_graph, layout = "manual", 
                           x = V(timed_graph)$x, y = V(timed_graph)$y) +
  geom_edge_link(aes(colour = color, 
                     edge_width = if_else(color == "#20854E", 1, 0.5))) +
  geom_edge_loop(aes(colour = color, 
                     edge_width = if_else(color == "#20854E", 1, 0.5))) +
  scale_edge_width(range = c(0.5, 1.5)) +
  scale_edge_color_manual(breaks = c("#0072B5", "#BC3C29", "#20854E"),
                          values = c("#0072B5", "#BC3C29", "#20854E"),
                          name = "Transmission link",
                          labels = c("SDSE", "S. pyogenes", "Shared")) +
  geom_scatterpie(aes(x = x, y = y, r = size * 1.5, group = name), 
                  cols = c("pyo", "sdse"),
                  data = igraph::as_data_frame(timed_graph, "vertices")) +
  scale_fill_manual(values = c("#BC3C29", "#0072B5")) +
  guides(fill = "none", edge_width = "none") +
  facet_wrap(~ factor(interval2, levels = c("Feb 2004", "Mar 2004", "Apr 2004", 
                                            "May 2004", "Jun 2004", 
                                            "Sep 2004", "Oct 2004", "Nov 2004",
                                            "Dec 2004", "Jan 2005", "Feb 2005",
                                            "Mar 2005", "Apr 2005", "May 2005",
                                            "Jun 2005")),
             nrow = 5, ncol = 3) + 
  coord_equal() + theme_bw(base_size = 20) +
  theme(axis.title = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.line = element_blank(),
        strip.background = element_rect(fill = "white"),
        panel.grid = element_blank())
```

```{r, echo=FALSE, fig.dim=c(10,15)}
transmission_net
```

Calculate number of unweighted transmission edges

```{r}
# All edges
nrow(union_all_df)
# SDSE unique
nrow(union_all_df %>% filter(color == "#0072B5"))
# pyo unique
nrow(union_all_df %>% filter(color == "#BC3C29"))
# Overlap edges
nrow(union_all_df %>% filter(color == "#20854E"))

# Within HH
# SDSE
nrow(union_all_df %>% filter(from == to, color %in% c("#0072B5", "#20854E")))
# pyo
nrow(union_all_df %>% filter(from == to, color %in% c("#BC3C29", "#20854E")))
```

Check transmission edges for putative throat isolate source

```{r}
# All edges
nrow(union_throat_df)
# SDSE unique
nrow(union_throat_df %>% filter(color == "#0072B5"))
# pyo unique
nrow(union_throat_df %>% filter(color == "#BC3C29"))
# Overlap edges
nrow(union_throat_df %>% filter(color == "#20854E"))

# Within HH and between HH edges
# SDSE
nrow(union_throat_df %>% filter(from == to, color %in% c("#0072B5", "#20854E")))
# pyo
nrow(union_throat_df %>% filter(from == to, color %in% c("#BC3C29", "#20854E")))
```


## Permutation test

Model of independent transmission is inferred using node label permutation to 
check if number of overlapping edges is what would be expected in the setting of 
independent transmission.

The publication uses 10,000 permutations. For speed, this markdown uses 1,000
permutations. This could be increased to 10,000 if reader would like to fully
replicate publication analysis.

```{r}
# Set number of iterations
no_iterations <- 1000

set.seed(1234)
permute_overlap <- tibble(iteration = 1:no_iterations,
                          overlap_edges = rep(0, no_iterations),
                          throat_overlap_edges = rep(0, no_iterations),
                          overlap_links = rep(0, no_iterations),
                          throat_overlap_links = rep(0, no_iterations))
sdse_perm_list <- list()
sdse_throat_perm_list <- list()

for (i in 1:no_iterations) {
  # Set permutation order which will then be consistent across all time points
  perm_order <- sample(rownames(sdse_time_matrix[[1]]), replace = FALSE)
  # Iterate through each time point and use same permutation order to maintain
  # transmission structure across time
  sdse_perm_iteration <- list()
  sdse_throat_perm_iteration <- list()
  for (j in 1:length(sdse_time_matrix)) {
    sdse_perm_iteration[[j]] <- sdse_time_matrix[[j]]
    # Permute nodes - maintain order in rows and columns so
    # within household transmission events remain within household
    rownames(sdse_perm_iteration[[j]]) <- colnames(sdse_perm_iteration[[j]]) <- 
      perm_order
    sdse_perm_graph <- graph_from_adjacency_matrix(sdse_perm_iteration[[j]],
                                                   mode="plus", weighted=TRUE, 
                                                   diag=TRUE)
    # Don't need to permute pyo as already permuted SDSE
    pyo_perm_graph <- graph_from_adjacency_matrix(pyo_time_matrix[[j]],
                                                  mode="plus", weighted=TRUE, 
                                                  diag=TRUE)
    sdse_throat_perm_iteration[[j]] <- sdse_time_matrix_carriage[[j]]
    # Permute nodes for throat matrix
    rownames(sdse_throat_perm_iteration[[j]]) <- 
      colnames(sdse_throat_perm_iteration[[j]]) <- perm_order
    sdse_throat_perm_graph <- graph_from_adjacency_matrix(
      sdse_throat_perm_iteration[[j]], mode="plus", weighted=TRUE, diag=TRUE)
    pyo_throat_perm_graph <- graph_from_adjacency_matrix(
      pyo_time_matrix_carriage[[j]], mode="plus", weighted=TRUE, diag=TRUE)
    # Calculate overlap with pyo using graph intersect
    intersect_graph <- sdse_perm_graph %s% pyo_perm_graph
    intersect_throat_graph <- sdse_throat_perm_graph %s% pyo_throat_perm_graph
    # Add to summary df
    # Links (weighted edges)
    permute_overlap[i, 4] <- permute_overlap[i, 4] + 
      sum(pmin(E(intersect_graph)$weight_1, E(intersect_graph)$weight_2))
    permute_overlap[i, 5] <- permute_overlap[i, 5] + 
      sum(pmin(E(intersect_throat_graph)$weight_1, 
               E(intersect_throat_graph)$weight_2))
    # Edges (unweighted)
    permute_overlap[i, 2] <- permute_overlap[i, 2] + length(E(intersect_graph))
    permute_overlap[i, 3] <- permute_overlap[i, 3] + 
      length(E(intersect_throat_graph))
  }
  sdse_perm_list[[i]] <- sdse_perm_iteration
}
```

Plot observed number of overlapping unweighted edges against permutation
distribution of the independent transmission model.

```{r, echo=FALSE}
permute_plot <- permute_overlap %>% ggplot(aes(x = overlap_edges)) + 
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = nrow(union_all_df %>% filter(color == "#20854E")), 
             linetype = "dashed", 
             colour = "firebrick", size = 1) +
  xlab("No. overlapping edges") + ylab("No. of permutations") +
  theme_bw(base_size = 20)

permute_plot
```

One sided p value of <= the observed value (i.e., hypothesis that there is
transmission interference).

```{r, echo=FALSE}
nrow(permute_overlap %>% 
       filter(overlap_edges <= 
                nrow(union_all_df %>% filter(color == "#20854E"))))/
  nrow(permute_overlap)
```
